|

### 2024-12-24 paper/manuscript.Rmd
drafting for the front-end.

---
title: |
  | The Promises and Pitfalls of Using Panel Data to Understand
  | Individual Belief Change
author:
  - Stephen Vaisey
  - Turgut Keskintürk
date: "`r format(Sys.time(), '%B %Y')`"
output: pdf_document
bibliography: ./helpers/references.bib
csl: ./helpers/asa.csl
abstract: "In this article, we investigate whether studies on political belief change can identify change trajectories at the individual level. Using simulations and empirical case studies, we propose a grid-search framework that allows researchers to evaluate the extent to which their target estimates generalize to individuals in their study population. We use simulated datasets to estimate plausible values for how many people changed, how much they changed, and who changed, based on observed survey response trajectories. Our results suggest that researchers should think carefully about the conditions under which they can make claims about belief change at the individual level."
fontsize: 11pt
fontfamily: mathpazo
geometry: margin=1in
figsintext: yes
header-includes:
- \usepackage{array}
- \usepackage{longtable}
- \usepackage{booktabs}
- \usepackage{float}
- \usepackage{caption}
- \usepackage{amsmath}
- \captionsetup{justification=raggedright,
                singlelinecheck=false}
---

@@

Studies on political change at the individual level---change in one's attitudes, values, or preferences---face a challenging problem: while our methods typically provide estimates for change and stability at the group level, our goal is often to understand these processes at the individual level.

This *group-to-person generalizability problem*, the claim that group-level findings may not generalize to each person [@mcmanus2023], has important implications for what we can and can't say about political change across the life-course. While the previous research quantified change with various methods---ranging from assessing the time order of longitudinal observations [@kiley2020; @vaisey2020] to multilevel decomposition of over-time variance [@lersch2023]---these studies have fallen short of disentangling whether they find a small amount of change in a large number of people, or a large amount of change in a small number of people.

This issue arises in a variety of ways in empirical practice, particularly when we want to theorize the mechanisms of political belief change. For example, the General Social Survey (GSS) shows that roughly 35% of the American adults in 2006 believed that "marijuana should be legalized." In less than 10 years, this support had increased to nearly 60%. As seen in Figure 1, this change occurred both *cross-sectionally*, where average support increased with time, and *longitudinally*, where at least some segments of the American population changed their minds on this policy.

@@

However, these two processes, cross-sectional change and longitudinal change, attract competing theories of political change, and group-level estimates may not properly adjudicate the underlying generative mechanisms. Following the same example, the cross-sectional change in beliefs about marijuana legalization may have occurred via two mechanisms: people may have changed their minds following a changing _zeitgeist_, or old generations who disapproved this policy might have been replaced by new generations approving it. Similarly, the longitudinal change could result from various mechanisms: large segments of the population may have shifted their opinions positively in small amounts, a small segment of the population may have made large changes, or a mix of changes---including both approvals and disapprovals---may have altered the overall balance.

While the cross-sectional processes received strong scholarly attention in the literature [@vaisey2016; @restrepoochoa2024; @bartels2014], questions of the second sort remained largely unexplored. However, given that theories of political ideology involve theoretical expectations about individual-level outcomes, the fact that we lack information about whether our models capture person-level empirical processes is detrimental to understanding adult belief change in the political context. Therefore, our article aims to address this gap by examining whether we can identify (a) *how many people changed*, (b) *how much they changed*, and (c) *who changed*, based on observed survey response trajectories of political beliefs in empirical data.

Using simulations and empirical case studies, we show that it is indeed possible to provide plausible scenarios as to which generative models are more likely to have produced empirical observations, even though a reliable identification of *who* changed occurs only under very narrow conditions. In doing this, we provide a simple grid-search procedure to aid researchers in evaluating the extent to which their target estimates may be generalized to the population of their study. We explore how competing generative models can lead to varying assumptions, emphasizing the importance for researchers to carefully consider the conditions under which claims about personal-level change can be made. Accordingly, we claim that when the meaning of "political change" is elusive and depends too much on how we define change in the first place, our empirical strategies could potentially mislead rather than inform, especially if we overlook the underlying generative mechanisms.

We start with an overview of the "latent variable model" of beliefs and the problems of reliability and resolution in survey response measurement [@alwin2007]. We then introduce our grid-search procedure in the context of three-wave panel data with binary outcomes, examining cases from the GSS on marijuana legalization and abortion. In the next section, we relax the period assumption and investigate how linear trajectories at the individual level can assist us in understanding underlying generative mechanisms in contexts with more than three waves. In the final empirical section, we analyze whether these generative models help us in classification tasks. We conclude by discussing the promises and pitfalls of panel data for understanding individual belief change.

@@

Alwin's [-@alwin2007] latent-variable approach to survey response is based on the assumption that there is a random variable $Y$, which captures a person $i$'s "true" position in a continuous latent construct. This random variable might stand in for any disposition, including one's political attitudes, values, and preferences, which allows us to define one's political position---at the level of one's "personal culture" [@lizardo2017]---as their *true score* $Y_i$.[^1]

[^1]: Naturally, the assumption that *there is indeed a true score* might miss cases where individuals hold contradictory ideas [@swidler1986] and are ambivalent about their true positions [@zaller1992; @zaller1992a]. In this article, we follow the latent variable model and assume that there *is* a true position for each individual.

The latent-variable approach reveals a fundamental problem in survey research: one's true score $Y_i$ is almost always realized with a certain degree of measurement error. Call this realization the *observed score* $y_i$. We often assume that $y_i$ represents the expected value of hypothetical draws from $Y_i$, with an $\epsilon$ term reflecting the measurement error in this process---i.e., $y_i = Y_i + \epsilon_i$. This error might result from various factors that introduce biases to how we extract information from individuals, given that problems like question construction and interviewer effects might contaminate how individuals reveal their true preferences [@alwin1991]. The inherent stochasticity in this measurement process implies that each draw from $Y_i$ might at least slightly differ from the true score, and this error term is difficult to circumvent with one measure, which is often the case in public opinion surveys.[^2] In this sense, there is often a degree of systematic gap between $Y_i$ and $y_i$.

[^2]: The idea that we should have repeated measures of $Y_i$ is the main intuition behind scale construction [@ansolabehere2008], with the explicit expectation that multiple measures can help us reduce measurement error.

Hence, the main problem is mapping one's true scores to their observed scores, and there are at least two ways we can fail in this context. First, our observed measures might have low *reliability*, such that a hypothetical set of repeated draws may significantly differ from one's true score, due to, e.g., poor question construction or individual idiosyncrasies. Second, survey response categories may be imperfect: the *resolution* of response categories may fail to distinguish consequential differences in people's true scores. In either case, the observed score may not reliably indicate one's true score.

## Separating True Change and Measurement Error in Panel Data

In the panel context, where we observe the same individuals across time, the measurement problem causes substantive questions: if true scores are imperfectly realized and we only have access to one's observed responses, how can we know if a person *actually changed* between time $t$ and $t + 1$, or whether the observed change is merely due to measurement error?

Scholars in public opinion research have attempted to address this issue in multiple ways, using structural equation models that explicitly quantify the measurement error component [@achen1975], strategies that integrate longitudinal surveys to exploit the information in the ordering of responses [@kiley2020], and various multilevel mixed-effects models that shrink responses based on population estimates to derive reliable trends [@lersch2023; @hout2016]. To some extent, all these strategies were fruitful to offset the problems associated with measurement error.[^3] That said, while these empirical strategies were largely successful in accounting for the amount of change *in the population*, they were rather silent about *who* changed in our observations.

[^3]: However, note that the use of repeated observations in the panel context comes with additional assumptions: it is rather straightforward to assume that multiple measures at time $t$ can be indicative of an underlying construct---e.g., we can average them to have an index---but the use of different time periods needs a *stability* assumption, that is, the reliability estimates may only be computed if one assumes that people do not really change between waves, which might confuse true change with measurement error [see, e.g., @hout2016].

Why is this important? After all, if we can produce reasonable estimates at the population level, this must surely be enough. We believe that there are at least two reasons why we should care about this problem. First, the inability to identify the individuals who truly change gives us a false sense of homogeneity. When we report an average change of, say, 0.10 standard deviations, we do not know whether this reflects uniform change in latent positions across all individuals, or significant changes in some while others remain stable. This is a big substantive difference, yet often overlooked in practice. Second, if we cannot identify the "changers" in the first place, our comparisons between groups---either across some pre-determined characteristic, or across treatment status---will average true processes with false negatives and false positives. This is obviously fine if the error occurs randomly, but it is challenging to verify this if individual-level accuracy is compromised. Hence, glossing over this problem discourages us to think about what exactly we are averaging.

## The Question of Individual Belief Change

@@

We pose these questions in a fixed-effects panel context [@allison2009; @vaisey2017], with varying individual trajectories across time [@rüttenauer2020]. This achieves two things. First, it allows us to shift the focus from *distributional change* to *individual change*, given that our theoretical interest is to understand how people change their political beliefs over the life-course. Second, it allows us to model response trajectories, rather than individual changes in response to a shock. Instead, our focus is "classification"---the extent to which we can discriminate those who change from those who do not change in a specific time window.

## A Grid-Search Procedure for Identifying Individual Belief Change

We propose a grid-search procedure to answer our research questions. The idea is straightforward:

(1) Generate simulated datasets with several parameters of interest---rate of change, strength of change, and the direction of change, so that we know the real data-generation process (DGP);
(2) Calculate the distribution of response trajectories in both the observed and simulated data;
(3) Compare the simulated trajectories to the observed trajectories, calculating a distance measure to quantify the extent to which specific simulated values reproduce the observed values;
(4) Replicate this analysis $N$ times for all possible parameter combinations;
(5) Inspect the parameter combinations that most closely approximate to the observed values.

This is simply what Gelman and Hill [-@gelman2007] call a "fake-data simulation," and it is powerful for at least two reasons. First, it allows us to produce reasonable estimates as to what plausible generative models might have produced the real data, ensuring that we can make educated guesses about the extent to which our parameter estimates might generalize to our populations. Second, it is generative, meaning that we can ask specific questions about response trajectories and modeling assumptions.

Our fake-data simulations involve several DGP components: a random variable $Y_{it}$ for true scores, with varying realizations $y_{it}$, representing a person $i$'s beliefs at time $t$; an indicator for change, $D_{it}$, operationalized as a non-reversible trigger if there is a respondent-level switch in true scores; an effect size, $\delta$, and a respondent level direction multiplier $\tau_i$ that indexes whether the change occurs in a positive or a negative direction. Using these components, the DGP follows

@@

Note that we allow actors to go through change at any time in the observation window of the study, meaning that $D_{it}$ indexes an actor's change status in their biographical history.

Since our research interest concerns individual trajectories of change, we specify three parameters for these DGPs: (1) *rate of change* refers to the proportion of our sample that goes through change in their political orientations, (2) *strength of change* refers to the standardized mean change in the true score when someone changes---described as Cohen's $d$---and (3) *direction of change* refers to the proportion of changers who change in positive and negative directions. The combination of these three parameters provides a complete picture of population movement in political beliefs.[^4]

[^4]: We define additional parameters that we fix in the case studies presented below: $N$ refers to the number of actors in the fake-data simulation, $T$ refers to the number of waves we have in the panel study, *balance* refers to the observed marginals in the real datasets, *reliability* refers to the item-reliability score regulating the extent to which a true score is realized in the survey context with given level of item reliability [@alwin2007], and *resolution* refers to the number of categories to simplify the latent variable, which is restricted to the binary case in this article.

Using these parameters, we can construct various generative models and compare the simulated trajectories to the observed trajectories to examine how closely certain parameter combinations approximate our empirical observations. In the next three sections, we will apply this procedure in different contexts. In all cases, we restrict our attention to latent variables with binary outcomes to simplify our analyses, but all results are generalizable to cases with higher resolution.[^repo]

[^repo]: The replication files for the article are stored in https://github.com/tkeskinturk/simulating_change

@@

# Study 2: Longer Panels and Linear Trajectories

@@

# Study 3: Detection of True Changers

@@

@@

In this article, we examined the identification of belief change in panel context. Using simulations and empirical case studies, we showed that a grid-search procedure that iterates over generative models can help us identify plausible data generation processes that may have produced our empirical observations, though the identification of individual cases occurs under narrow conditions. We argued that the *group-to-person generalizability problem* is an impediment to a robust understanding of political belief change, and scholars should understand the extent to which their target estimates can generalize to the population of their study. As such, we proposed a simulation approach that takes response generation seriously, which may help us move forward in addressing this problem.

This article has several implications for the study of belief change across the life-course. First, our findings suggest that recent debates about the existence or non-existence of belief change in adult populations [@kiley2020; @lersch2023] should tackle the heterogeneous nature of change and stability in individual preferences since group-level measures may hamper our understanding of individual mechanisms. Instead, scholars need to specify the generative models that aggregate average estimates imply. Analyses that examine population patterns over time allow us to adjudicate whether compositional changes or individual changes drive cumulative political change [@bartels2014; @vaisey2016]. Similarly, we need robust and reliable panel studies to decompose whether individual changes occur across all individuals---implying period changes---or in specific groups of people---implying political heterogeneity.

>>>>

---
title: |
  | The Promises and Pitfalls of Using Panel Data to Understand Individual Belief Change[^thanks]
author:
  - Turgut Keskintürk[^TK]
  - Pablo Bello[^PB]
  - Stephen Vaisey[^SV]
date: "`r format(Sys.time(), '%B %Y')`"
output: pdf_document
bibliography: ./helpers/references.bib
csl: ./helpers/asa.csl
abstract: "We investigate whether studies on political belief change can identify change trajectories at the individual level. Using simulations and case studies, we propose a grid-search framework that allows researchers to evaluate the extent to which their target estimates generalize to individuals in their study population. We use simulated datasets to estimate plausible values for how many people changed, how much they changed, and who changed, based on observed survey response trajectories. Our results suggest that researchers should think carefully about the conditions under which they may make claims about belief change at the individual level. We provide an R package to help researchers implement this procedure in their own work."
fontsize: 11pt
fontfamily: mathpazo
geometry: margin=1in
figsintext: yes
header-includes:
- \usepackage{array}
- \usepackage{float}
- \usepackage{caption}
- \usepackage{amsmath}
- \captionsetup{justification=raggedright,
                singlelinecheck=false}
---

[^thanks]: We thank the participants of the *Worldview Lab at Kenan Institute for Ethics* at Duke University for their insightful and helpful comments on this project.
[^TK]: Department of Sociology, Duke University, `turgut.keskinturk@duke.edu`.
[^PB]: Department of Sociology, Duke University, `pablo.bello@duke.edu`.
[^SV]: Department of Sociology, Duke University, `stephen.vaisey@duke.edu`.

@@

Studies on political change at the individual level---change in one's beliefs and preferences---face a challenging problem: while our methods typically provide estimates for change and stability at the group level, our goal is often to understand these processes at the individual level.

This *group-to-person generalizability problem*, the claim that group-level findings may not generalize to each person [@mcmanus2023], has important implications for what we can and can't say about political change across the life-course. While the previous research quantified change with various methods---ranging from assessing the time order of longitudinal observations [@kiley2020; @vaisey2020] to multilevel decomposition of over-time variance [@lersch2023]---these studies have fallen short of disentangling whether they find a small amount of change in a large number of people, or a large amount of change in a small number of people. In other words, while social science researchers have become widely sophisticated in evaluating the average levels of change over time in a population, they have mostly overlooked the fact that these average changes are produced by a combination of various compositional processes.

This issue arises in a variety of ways in empirical practice, particularly when we want to theorize the mechanisms of political belief change. For example, the General Social Survey (GSS) shows that roughly 35% of the American adults in 2006 believed that "marijuana should be legalized." In less than 10 years, this support had increased to nearly 60%. As seen in Figure 1, this change occurred both *cross-sectionally*, where average support increased with time, and *longitudinally*, where at least some segments of the American population changed their minds on this policy over these periods.

@@

However, these two processes, cross-sectional change and longitudinal change, attract competing theories of political change, and group-level estimates may not properly adjudicate the underlying data generation processes. Following the same example, the cross-sectional change in beliefs about marijuana legalization may have occurred via two processes: people may have changed their minds following a changing _zeitgeist_, or old generations who disapproved this policy might have been replaced by new generations approving it. Similarly, the longitudinal change could result from various processes: large segments of the population may have shifted their opinions positively in small amounts, a small segment of the population may have made large changes, or a mix of changes---including both approvals and disapprovals---may have altered the overall balance.

While the cross-sectional processes have received strong scholarly attention in the literature [@vaisey2016; @restrepoochoa2024; @bartels2014], questions of the second sort remained largely unexplored. However, given that theories of political ideology involve theoretical expectations about individual-level outcomes, the fact that we lack information about whether our models capture person-level empirical processes is detrimental to understanding adult belief change in the political context. Therefore, our article aims to address this gap by examining whether we can identify (a) *how many people changed*, (b) *how much they changed*, and (c) *who changed*, based on observed survey response trajectories of political beliefs in empirical data.

Using simulations and empirical case studies, we show that it is indeed possible to provide plausible scenarios as to which generative models are more likely to have produced empirical observations, even though a reliable identification of *who* changed occurs only under very narrow conditions. In doing this, we build on the "Approximate Bayesian Computation" (APC) literature [@beaumont2010] and provide a simple grid-search procedure to help social researchers in evaluating the extent to which their target estimates might be generalized to the population of their studies. Using this procedure, we show that it is possible to approximate to the underlying data generation processes for *rate of change*, *strength of change*, and *direction of change* in a panel data. We provide an R package, `gridsearch`, to help researchers in implementing this procedure in their own work.

We emphasize a striking heterogeneity in these underlying processes even for a seemingly straightforward case like the marijuana legalization in the GSS, showing that theoretically opposite claims---e.g., large swaths of the population increasing their support for the policy, as opposed to a small group of people completely switching their positions---may very well account for the same data. This finding is particularly important in the longitudinal panel context, which, for most applications, represents the gold standard in evaluating change in observational social research. As such, we claim that when the meaning of "political change" is elusive, and depends too much on how we define change in the first place, our current empirical strategies could potentially mislead rather than inform, especially if we overlook the underlying generative mechanisms.

We start with a general overview of the "latent variable model" of beliefs, the problems of reliability and resolution in survey measurement, and the compositional sources of aggregate change. We then introduce our grid-search procedure in the context of three-wave panel data with binary outcomes. Next, we relax this assumption and investigate how linear trajectories at the individual level can assist us in contexts with more than three waves. In the final section, we analyze whether this procedure helps us in classification tasks, particularly in evaluating *who* changed. We conclude by discussing the promises and pitfalls of panel data for understanding individual political change.

@@

The latent-variable approach to survey response is based on the assumption that there is a random variable $Y$, which captures a person $i$'s "true" position in a continuous latent distribution [@alwin2007]. This random variable might stand in for any disposition, including one's political beliefs and preferences, which allows us to define one's political position as their *true score* $Y_i$.[^truescore]

[^truescore]: Naturally, the assumption that *there is indeed a true score* might miss cases where individuals hold contradictory ideas [@swidler1986] and are ambivalent about their true positions [@zaller1992; @zaller1992a]. In this article, we follow the latent variable model and assume that there *is* a true position for each individual; hence, bracketing the cases where this ambivalence needs to be incorporated into the survey response measurement model.

The latent-variable approach reveals a fundamental problem in survey research: one's true score $Y_i$ is almost always realized with a certain degree of measurement error. Call this realization the *observed score* $y_i$. We often assume that $y_i$ represents the expected value of hypothetical draws from $Y_i$, with an $\epsilon$ term reflecting the measurement error in this process---i.e., $y_i = Y_i + \epsilon_i$. This error might result from various factors that introduce biases to how we extract information from individuals, given that problems like question construction and interviewer effects might contaminate how individuals reveal their true preferences [@alwin1991]. The inherent stochasticity in this measurement process implies that each draw from $Y_i$ might at least slightly differ from the true score, and this error term is difficult to circumvent with one measure, which is often the case in public opinion surveys.[^scales] In this sense, there is often a degree of systematic gap between $Y_i$ and $y_i$.

[^scales]: The idea that we should have repeated measures of $Y_i$ is the main intuition behind scale construction [@ansolabehere2008], with the explicit expectation that multiple measures can help us reduce measurement error.

In the panel context, where we observe the same people multiple times, this measurement problem has substantive implications: if true scores are imperfectly realized and we only have surface access to a person's observed survey responses, how can we know if they *actually* changed between time $t$ and $t + 1$, or whether the observed change is merely due to measurement error?

Scholars in public opinion research have attempted to address this issue in multiple ways, using structural equation models that explicitly quantify the measurement error component [@achen1975], strategies that integrate longitudinal surveys to exploit the information in the ordering of responses [@kiley2020], and various multilevel mixed-effects models that shrink responses based on population estimates to derive reliable trends [@lersch2023; @hout2016]. To some extent, all these strategies were fruitful to offset the problems associated with measurement error.[^stability] That said, while these empirical strategies were largely successful in accounting for the amount of change *in the population*, they were rather silent about changes *at the level of individuals*.

[^stability]: However, note that the use of repeated observations in the panel context comes with additional assumptions: it is rather straightforward to assume that multiple measures at time $t$ can be indicative of an underlying construct---e.g., we can average them to have an index---but the use of different time periods needs a *stability* assumption, that is, the reliability estimates may only be computed if one assumes that people do not really change between waves, which might confuse true change with measurement error [see, e.g., @hout2016].

## The Compositional Sources of Aggregate Change

Suppose we have a panel data, where individuals $i = 1, ..., N$ are observed over two time periods $t_0$ and $t_1$. Let $y_{it}$ represent an observed survey response. To see how much observed change there is at the individual level, we can simply define a change quantity, $\delta y_{i, t_0 \to t_1} = y_{it_1} - y_{it_0}$, the average of which, $\bar{y}$, represents the mean change in the dataset from $t_0$ to $t_1$.

This quantity is a natural candidate for describing aggregate belief change, but, under the hood, it summarizes three compositional processes in one metric. Let us define these processes intuitively. The first component is $R$, representing the proportion of individuals in our dataset with $\delta y_{i, t_0 \to t_1} \neq 0$, i.e., the proportion of individuals who changed in this period. Call this *the rate of change*. The second component is $S$, representing the "average" magnitude of change among those who changed.[^heterogeneity] Call this *the strength of change*. The third component is $D$, representing the net directionality of change, i.e., the difference in proportions of positive and negative changers in our observations. Call this *the direction of change*. If we combine these three components, we can represent $\bar{y}$ simply as

$$
\bar{y} = R \times S \times D
$$

[^heterogeneity]: Of course, we might further decompose this average strength to a distribution of heterogeneous change trajectories across individuals. For analytical and expository clarity, we bracket this issue from the scope of this article.

This simple decomposition should let us appreciate that the average change score may be generated via multiple data generation processes, each having strongly different theoretical implications.

For instance, if we want to evaluate whether a society became more pro-redistribution over time, we need to note that an increase in pro-redistributive preferences can occur via all or most people reporting higher preference for redistribution, some people switching their positions altogether while some others remain completely stable, or everyone changing in one or the other direction, and the net score we observe is simply the difference in their overall trajectories. The same reasoning applies to the GSS's marijuana legalization policy that we have discussed before. Most discussions in political belief change---confidence in legislative groups or financial institutions, beliefs about the economy, or ethnic attitudes---have the exact same question about their data generation processes. Simply put, the findings about change at the population level may simply result from a variety of conflicting, though observationally equivalent, data generation processes.

We can now complicate this composition question by noting, once again, the survey measurement problem of mapping one's true scores to their observed scores, which often fails in at least two ways: (1) our observed measures might have low *reliability*, such that a hypothetical set of repeated draws may significantly differ from one's true score, due to, e.g., poor question construction or individual idiosyncrasies, and (2) question response categories may be imperfect: the *resolution* of response categories may fail to distinguish consequential differences in true scores. Therefore, the average change captured in $\bar{y}$ may represent a biased or imperfect estimate of change to begin with.

@@

We pose these questions in a fixed-effects longitudinal panel data context [@allison2009; @vaisey2017], with varying individual trajectories across time periods [@rüttenauer2020]. This allows us to shift the focus from *aggregate change* to *individual change*, given that our theoretical interest is to understand how people change their political beliefs over the life-course.

## A Grid-Search Procedure

We build on the Approximate Bayesian Computation (APC) literature [@beaumont2010] and propose a grid-search procedure to answer our research questions. The APC approach involves simulating parameter values given the data and evaluating the distribution of these parameters that potentially generated our observations. We build a rejection sampler algorithm that

1) Simulates data with several parameters of interest for our DGPs---rate of change, strength of change, direction of change, and reliability of response measurement using

@@

    where we have a random variable $Y_{i}$ for true scores, with varying realizations $y_{it}$, representing a person $i$'s beliefs at time $t$; an indicator for change, $D_{it}$, operationalized as a non-reversible trigger if there is a respondent-level change in true scores; an effect size, $\delta$, and a respondent level direction multiplier $\tau_i$ that indexes whether the change is positive or negative.[^changing]

2) Calculates the distribution of response trajectories in both the observed and simulated data, by either tabulating response patterns across time or estimating slope coefficients;

3) Compares the simulated trajectories to the observed trajectories, calculating a "distance" score to quantify the extent to which specific simulated values reproduce the observed values;

[^changing]: Note that we "allow" actors to go through change at any time in the observation window of the study, meaning that $D_{it}$ is variable over $t$ precisely because actors can change any time in their biographical history.

4) Rejects the simulated data if the distance measure is above a certain pre-specified "caliper" value, and accepts it otherwise;

5) Replicates this analysis $N$ times by drawing samples from the specified priors;

6) Inspects the parameter combinations that most closely approximate to the observed values.

The grid-search approach we propose is powerful for at least two reasons. First, it allows us to produce reasonable estimates as to what plausible generative models might have produced the real data, ensuring that we can make educated inferences about the extent to which our parameter estimates might generalize to our populations. Second, it is generative, meaning that we can ask specific questions about response trajectories and modeling assumptions.[^parameters]

[^parameters]: We define additional parameters that we fix in the case studies presented below: $N$ refers to the number of actors in the fake-data simulation, $T$ refers to the number of waves we have in the panel study, *balance* refers to the observed marginals in the real datasets, allowing us to "cut" the continuous distribution to 0s and 1s, and *resolution* refers to the number of categories to simplify the latent variable, which is restricted to the binary case in this article.

In the next four sections, we apply this procedure in different contexts. In all cases, we restrict our attention to latent variables with binary outcomes to simplify our analyses. The replication files for the article are stored in https://osf.io/AAAAA. The R package for the use of this approach, `gridsearch`, can be accessed via the replication repo or https://tkeskinturk.github.io/gridsearch/.

\newpage

@@

## The Case of Generalized Trust

# Study 2: Simulation Studies and Performance Assessment

# Study 3: Longer Panels and Linear Trajectories

@@

# Study 4: Detection of True Changers

@@

\clearpage

@@

In this article, we examined the identification of belief change in panel context. Using simulations and empirical case studies, we showed that a grid-search procedure that iterates over generative models can help us identify plausible data generation processes that may have produced our empirical observations, though the identification of individual cases occurs under narrow conditions. We argued that the *group-to-person generalizability problem* is an impediment to a robust understanding of political belief change, and scholars should understand the extent to which their target estimates may generalize to the population of their study. As such, we proposed a simulation approach that takes response generation seriously, which may help us move forward in addressing this problem.

This article has several implications for the study of political belief change across the life-course. First, our findings suggest that recent debates about the existence or non-existence of belief change in adult populations [@kiley2020; @lersch2023] should tackle the heterogeneous nature of change and stability in individual preferences since group-level measures may hamper our understanding of individual mechanisms. Instead, scholars need to specify the generative models that aggregate average estimates imply. Analyses that examine population patterns over time allow us to adjudicate whether compositional changes or individual changes drive cumulative political change [@bartels2014; @vaisey2016]. Similarly, we need robust and reliable panel studies to decompose whether individual changes occur across all individuals---implying period changes---or in specific groups of people---implying political heterogeneity.

### 2024-12-24 paper/revision_memo.Rmd
drafting for the front-end.

- We added a new item using the *British Elections Study*'s panel study, generalized trust, an item collected between 2019 and 2024 across 6 years and 5 waves, in a potentially disruptive time period due to COVID-19. As a more general note, we believe that the tool we propose should apply in all cases, but we agree that a more thorough set of examples are helpful.
- Additionally, we now present a simulation study that applies our protocol in different DGPs and assesses the conditions under which we can make better guesses about the true DGPs (see below for more details). We think this new section provides important answers about the questions of generalizability, panel span, varying DGP compositions, and related issues.

@@

## The Clarity of the Grid Search Discussions

@@

We implemented three solutions for this issue. First, we now introduce our grid-search algorithm as an R package, `gridsearch`, which will accompany to this article. As we detail in the paper, we revised our procedure and substantially increased its efficiency, which means that it will be easy for people to implement our procedure in their own projects. Second, we now include a full replication package for the analyses presented in this project as a footnote in the article. We were planning to do so along with publication, but for a paper of this kind, it is beneficial for the reviewers to see the full code, so we attached a view-only (anonymized) link in the manuscript. Finally, we made substantive rewrites to the manuscript to make our protocol as clear and concise as possible. We hope that the R package, the replication repo, and these rewrites make our strategy more clear.

@@

>>>>

- We added a new item using the *British Elections Study*'s panel study, generalized trust, an item collected between 2019 and 2024 across 6 years and 5 waves, in a potentially disruptive time period due to COVID-19. As a more general note, we believe that the tool we propose should apply in all cases, but we agree that a more thorough set of examples is helpful.
- Additionally, we now present a simulation study that applies our protocol in different DGPs and assesses the conditions under which we can make better guesses about the true DGPs (see below for more details). We think this new section provides important answers about the questions of generalizability, panel span, varying DGP compositions, and related issues.

@@

## The Grid-Search Procedure

@@

We implemented three solutions and a larger revision for this problem:

1. We updated the mechanics of our approach by using Approximate Bayesian Computation [@beaumont2010], which increased the efficiency and generalizability of our procedure.
2. We now introduce our grid-search algorithm as an accompanying R package, `gridsearch`. As we detailed in the paper, we revised our procedure and substantially increased its efficiency, which means that it will be easy for people to implement our method in their own projects.
3. We included a full (anonymized) replication package for the analyses presented in this project.
4. We made substantial rewrites to the manuscript to increase the clarity of our protocol.

We hope that the R package, the replication repo, and these rewrites make our strategy more clear.

@@

# References
